### 2018-05-22 Paper/examples.tex
..

We consider the following semi-linear elliptic PDE on a 2D square domain ([0,1]$\times$[0,1]):

\begin{equation}\label{eq:semilinear}
\begin{aligned}
-\kappa \Delta u + c u^3 &= q, \quad \text{in } \Omega\\
 u &= 0. \quad \text{on } \partial \Omega
\end{aligned}
\end{equation}
%
where $u$ is referred to as the velocity, $\kappa$ and $c$ are coefficients of the diffusion term
and the non-linear term in the above equation respectively. The velocity, $u$ is random since the
source term, $q$ in RHS is modelled as a truncated summation with random coefficients:

\be
q = \sum\limits_{i=1}^{N=8}\alpha_i\sin\left(\frac{i\pi x}{8}\right)\cos\left(\frac{i\pi y}{8}\right)
\label{eq:source}
\ee
%
Hence, $u~=~u(\vec{\theta})$ with $\vec{\theta} = (\kappa,c,\alpha_1,\alpha_2,\ldots,\alpha_{8})^T$, a vector
of uncertain parameters. Table~\ref{tab:elliptic} provides distributions of the uncertain
input parameters. The solution of~\eqref{eq:semilinear} for
a particular set of values of the uncertain parameters is illustrated below in Figure~\ref{fig:usoln}.

\begin{table}[htbp]\label{tab:elliptic}
\renewcommand*{\arraystretch}{1.2}
\begin{center}
\begin{tabular}{cl}
\toprule
\textbf{Parameter} & \textbf{Distribution} \\ 
\bottomrule
$\kappa$ & $\mathcal{U}$[0.05,0.1] \\
$c$ & $\mathcal{U}$[1.0,2.0] \\
$\alpha_i$ & $\mathcal{U}$[0.0,4.0] \\
\bottomrule
\end{tabular}
\end{center}
\caption{Description and distributions of uncertain parameters in the 
semi-linear elliptic PDE in~\eqref{eq:semilinear}.}
\end{table}

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.6\textwidth]{./Figures/u_soln}
  \caption{Plotted above is the solution, $u(x,y)$ of~\eqref{eq:semilinear} at 
  of $\kappa$~=~0.075, $c$~=~1.5, and $\alpha_i$~=~4.0.}
\label{fig:usoln}
\end{center}
\end{figure}

@@

\be
\mathcal{F}(\kappa, c, \theta) = \frac{1}{|D|} \int_D u(x; \kappa, c, \theta) \, dx, 
\label{eq:qoi}
\ee
%
where $D$ is the region: $[2/5, 3/5] \times [2/5, 3/5]$. Once again, the availability of
cheap model evaluations allow us to obtain converged estimates of the Sobol
total-effect index ($\mathcal{T}(\theta_i)$), plotted in Figure~\ref{fig:sense_elliptic}.

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.46\textwidth]{./Figures/sense_elliptic}
  \includegraphics[width=0.48\textwidth]{./Figures/ub_conv_elliptic}
\caption{
Left: Sobol' total sensitivity index, $\mathcal{T}(\theta_i)$ for uncertain parameters in the 
semi-linear elliptic PDE~\eqref{eq:semilinear}. Right: 
Estimates of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$) for each uncertain parameter,
obtained using $N$ = 5, 10, 15, and 20 samples in the full parameter space.}
\label{fig:sense_elliptic}
\end{center}
\end{figure}
%
Sensitivity predictions based on the screening metric, $\widehat{\mathcal{C}_i\mu_i}$ are found to
be in close agreement with $\mathcal{T}(\theta_i)$, even for the case when $N$ = 5. Moreover, 
the uncertainty in the QoI in \eqref{eq:qoi} is observed to be largely dependent on $c$, 
$\alpha_2$, $\alpha_3$, $\alpha_4$, and $\alpha_5$. These observations underscore the
potential for computational gains by constructing an ROS in the 5D parameter space. We 
illustrate the comparison of convergence characteristics of the PCEs constructed in the
full parameter space (10D) and the reduced space (5D) in Figure~\ref{fig:conv_elliptic} (left). 
As expected, the ROS converges at a much faster rate. Using model evaluations at 90
sample points, $\epsilon_{\mbox{\tiny{LOO}}}$ is found to be two orders of magnitude
smaller than that in the case of full-surrogate ($\mathcal{O}(10^{-4}$) vs. $\mathcal{O}(10^{-2}$)). 
Consequently, the computational effort for constructing the ROS in the present test problem
is expected to be much smaller.

@@

>>>>

We consider the following semi-linear elliptic PDE on a 2D square domain ([0,1]$\times$[0,1])
with Dirichlet boundary conditions:

\begin{equation}\label{eq:semilinear}
\begin{aligned}
-\kappa \Delta u + c u^3 &= q, \quad \text{in } \Omega\\
 u &= 0. \quad \text{on } \partial \Omega
\end{aligned}
\end{equation}
%
where $u$ is referred to as the velocity, $\kappa$ and $c$ are coefficients of the diffusion term
and the non-linear term in the above equation respectively. The velocity, $u$ is random since the
source term, $q$ in the RHS is modelled as a truncated summation with random coefficients:

\be
q = \sum\limits_{i=1}^{N=8}\alpha_i\sin\left(\frac{i\pi x}{8}\right)\cos\left(\frac{i\pi y}{8}\right)
\label{eq:source}
\ee
%
Hence, $u~=~u(\vec{\theta})$ with $\vec{\theta} = 
(\kappa,c,\alpha_1,\alpha_2,\ldots,\alpha_{8})^T$, a vector
of uncertain parameters. Distributions of the uncertain
input parameters are tabulated below in Figure~\ref{fig:elliptic}~(left).
The solution of~\eqref{eq:semilinear} for
a fixed set of values of the uncertain parameters is also 
illustrated.

\begin{figure}[htbp]
\begin{center}
\begin{minipage}[htbp]{.25\linewidth}
\vspace{0pt}
%\centering
\hspace{-25mm}
\begin{tabular}{cl}
\toprule
$\theta_i$ & \textbf{Distribution} \\ 
\bottomrule
$\kappa$ & $\mathcal{U}$[0.05,0.1] \\
$c$ & $\mathcal{U}$[1.0,2.0] \\
$\alpha_i$ & $\mathcal{U}$[0.0,4.0] \\
\bottomrule
\end{tabular}
\end{minipage}
\hspace{-20mm}
\begin{minipage}[htbp]{.25\linewidth}
\vspace{0pt}
%\centering
\includegraphics[width=3.0in]{./Figures/u_soln.png}
\end{minipage}%
\end{center} 
\caption{Left: Table providing distributions of the individual
uncertain parameters in~\eqref{eq:semilinear}. Right: Solution
of the 2D semi-linear elliptic PDE~\eqref{eq:semilinear} using
$\kappa$~=~0.075, $c$~=~1.5, and $\alpha_i$~=~4.0}
\label{fig:elliptic}
\end{figure}

%\begin{table}[htbp]
%\begin{tabular}{cc}
%
%  \includegraphics[width=0.45\textwidth]{./Figures/u_soln.png} &
%\begin{tabular}{cl}
%\toprule
%$\theta_i$ & \textbf{Distribution} \\ 
%\bottomrule
%$\kappa$ & $\mathcal{U}$[0.05,0.1] \\
%$c$ & $\mathcal{U}$[1.0,2.0] \\
%$\alpha_i$ & $\mathcal{U}$[0.0,4.0] \\
%\bottomrule
%\end{tabular}
%\\
%\end{tabular}
%\end{table}

@@

\be
\mathcal{F}(\kappa, c, \theta) = \frac{1}{|D|} \int_D u(x; \kappa, c, \theta) \, dx, 
\label{eq:qoi}
\ee
%
where $D$ is the region: $[2/5, 3/5] \times [2/5, 3/5]$ in $\mathbb{R}^{x\times y}$. 
The availability of
cheap model evaluations allow us to obtain converged estimates of the Sobol
total-effect index ($\mathcal{T}(\theta_i)$), plotted in Figure~\ref{fig:sense_elliptic}~(left),
with reasonable effort.

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.46\textwidth]{./Figures/sense_elliptic}
  \includegraphics[width=0.48\textwidth]{./Figures/ub_conv_elliptic}
\caption{
Left: Sobol' total sensitivity index, $\mathcal{T}(\theta_i)$ for uncertain parameters in the 
semi-linear elliptic PDE~\eqref{eq:semilinear}. Right: 
Estimates of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$) for each uncertain parameter,
obtained using $N$ = 5, 10, 15, and 20 samples in the full parameter space.}
\label{fig:sense_elliptic}
\end{center}
\end{figure}
%
Sensitivity predictions based on the screening metric, $\widehat{\mathcal{C}_i\mu_i}$,
plotted in Figure~\ref{fig:sense_elliptic}~(right), are found to
be in close agreement with $\mathcal{T}(\theta_i)$, even for the case when $N$ = 5. As $N$
is increased from 5 to 20, estimates of the screening metric are observed to converge.
Based on the trends observed in Figure~\ref{fig:sense_elliptic}, it can be said that
the uncertainty in the QoI in \eqref{eq:qoi} is largely dependent on $c$, 
$\alpha_2$, $\alpha_3$, $\alpha_4$, and $\alpha_5$. These observations underscore the
potential for computational gains by constructing an ROS in the 5D parameter space. We 
illustrate the comparison of convergence characteristics of the PCEs constructed in the
full parameter space (10D) and the reduced space (5D) in Figure~\ref{fig:conv_elliptic} (left). 
As expected, the ROS converges at a much faster rate. Using model evaluations at 90
sample points, $\epsilon_{\mbox{\tiny{LOO}}}$ is found to be two orders of magnitude
smaller than that in the case of full-surrogate ($\mathcal{O}(10^{-4}$) vs. $\mathcal{O}(10^{-2}$)). 
Consequently, the computational effort for constructing the ROS in the present test problem
is expected to be much smaller.

@@

Once again, we verify the accuracy of the ROS by estimating $\epsilon_{\mbox{\tiny{L-2}}}$
using model evaluations at 100 independent Monte-Carlo samples in the 10D parameter
space. The L-2 norm was estimated to be 0.05 i.e. the ROS was found to be accurate
within 5$\%$. Additionally, we compare PDFs of the QoI, $\mathcal{F}(\kappa,c,\theta)$,
generated using the full-space and the reduced-space surrogates in 
Figure~\ref{fig:conv_elliptic}~(right). The two PDFs are observed to be in excellent 
agrement with each other. Hence, the ROS is successfully verified and could be
used in lieu of the full-space surrogate to reduce computational effort.
